---
title: "STA302 Final Project"
author: "Qiannan Zhao"
output: pdf_document
---

```{r read data and split the dataset}
bit <- read.csv(file="FinalDataset_Bit.csv", header=T)
str(bit)
# create a 50/50 split in the data
set.seed(1)
train <- bit[sample(1:nrow(bit),373, replace=F), ]
test <- bit[which(!(bit$Time %in% train$Time)),]

```


```{r}
mtr <- apply(train[,-c(1)], 2, mean)
sdtr <- apply(train[,-c(1)], 2, sd)

mtest <- apply(test[,-c(1)], 2, mean)
sdtest <- apply(test[,-c(1)], 2, sd)

```

We can take these and add them nicely to a table:

Variable | mean (s.d.) in training | mean (s.d.) in test
---------|-------------------------|--------------------
`r names(test)[3]` | `r round(mtr[1], 3)` (`r round(sdtr[1], 3)`) | `r round(mtest[1], 3)` (`r round(sdtest[1], 3)`)
`r names(test)[4]` | `r round(mtr[2],3)` (`r round(sdtr[2],3)`) | `r round(mtest[2],3)` (`r round(sdtest[2],3)`)
`r names(test)[5]` | `r round(mtr[3],3)` (`r round(sdtr[3],3)`) | `r round(mtest[3],3)` (`r round(sdtest[3],3)`)
`r names(test)[6]` | `r round(mtr[4],3)` (`r round(sdtr[4],3)`) | `r round(mtest[4],3)` (`r round(sdtest[4],3)`)
`r names(test)[7]` | `r round(mtr[5],3)` (`r round(sdtr[5],3)`) | `r round(mtest[5],3)` (`r round(sdtest[5],3)`)
`r names(test)[8]` | `r round(mtr[6],3)` (`r round(sdtr[6],3)`) | `r round(mtest[6],3)` (`r round(sdtest[6],3)`)
`r names(test)[9]` | `r round(mtr[7],3)` (`r round(sdtr[7],3)`) | `r round(mtest[7],3)` (`r round(sdtest[7],3)`)
`r names(test)[10]` | `r round(mtr[8],3)` (`r round(sdtr[8],3)`) | `r round(mtest[8],3)` (`r round(sdtest[8],3)`)
`r names(test)[11]` | `r round(mtr[9],3)` (`r round(sdtr[9],3)`) | `r round(mtest[9],3)` (`r round(sdtest[9],3)`)
`r names(test)[12]` | `r round(mtr[10],3)` (`r round(sdtr[10],3)`) | `r round(mtest[10],3)` (`r round(sdtest[10],3)`)

Table: Summary statistics in training and test dataset, each of size 373.


```{r}
full <- lm(BTC_Price ~ ., data=train[,-c(1)])
summary(full)

# check conditions for checking model assumptions
pairs(train[,-c(1,2,3,4,5,6,7)])
plot(train$BTC_Price ~ fitted(full), main="Y vs Fitted", xlab="Fitted", ylab="BTC_Price")
lines(lowess(train$BTC_Price ~ fitted(full)), lty=2)
abline(a = 0, b = 1)
```

```{r}
# check model assumptions
par(mfrow=c(2,2))
plot(rstandard(full)~fitted(full), xlab="fitted", ylab="Residuals")
for(i in c(2:10)){
  plot(rstandard(full)~train[,i], xlab=names(train)[i], ylab="Residuals")
}

qqnorm(rstandard(full))
qqline(rstandard(full))

# try transformations
library(car)
p <- powerTransform(cbind(bit[,-c(1)]))
summary(p)

vif(full)
```

```{r}
# so transform just the response and Driving Accuracy in both the train and test
train$logBTC_Price <- log(train$BTC_Price)
train$logETH._Price <- log(train$ETH._Price)
train$logRipple_Price <- log(train$Ripple_Price)
train$logMiner.s_Revenue <- log(train$Miner.s_Revenue)
train$logGold_Price <- log(train$Gold_Price)
train$logMSCI_World <- log(train$MSCI_World)
train$logAverage_Block_Size <- log(train$Average_Block_Size)

test$logBTC_Price <- log(test$BTC_Price)
test$logETH._Price <- log(test$ETH._Price)
test$logRipple_Price <- log(test$Ripple_Price)
test$logMiner.s_Revenue <- log(test$Miner.s_Revenue)
test$logGold_Price <- log(test$Gold_Price)
test$logMSCI_World <- log(test$MSCI_World)
test$logAverage_Block_Size <- log(test$Average_Block_Size)

#construct the transformed model and check assumptions
full2 <- lm(logBTC_Price ~ ., data=train[,-c(1,2,3,4,5,8,9,10)])
summary(full2)
pairs(train[,-c(1,2,3,4,5,6)])
plot(train$BTC_Price ~ fitted(full2), main="Y vs Fitted", xlab="Fitted", ylab="BTC_Price")
lines(lowess(train$BTC_Price ~ fitted(full2)), lty=2)
abline(a = 0, b = 1)

par(mfrow=c(2,2))
plot(rstandard(full2)~fitted(full2), xlab="fitted", ylab="Residuals")
for(i in c(2:10)){
  plot(rstandard(full2)~train[,i], xlab=names(train)[i], ylab="Residuals")
}

qqnorm(rstandard(full2))
qqline(rstandard(full2))


vif(full2)
```


```{r model3}
mod3 <- lm(logBTC_Price ~ logETH._Price + logRipple_Price + logMiner.s_Revenue + logGold_Price + logMSCI_World + logAverage_Block_Size, data=train)
summary(mod3)

vif(mod3)
which(cooks.distance(mod3)>qf(0.5, 7, 373-7))
which(abs(dffits(mod3)) > 2*sqrt(8/373))
par(mfrow=c(1,1))
plot(rstandard(mod3)~train$logETH._Price)
plot(rstandard(mod3)~train$logRipple_Price)
plot(rstandard(mod3)~train$logMiner.s_Revenue)
plot(rstandard(mod3)~train$logGold_Price)
plot(rstandard(mod3)~train$logMSCI_World)
plot(rstandard(mod3)~train$Average_Block_Size)
qqnorm(rstandard(mod3))
qqline(rstandard(mod3))


```
```{r model4}
mod4 <- lm(logBTC_Price ~ logGold_Price + logMiner.s_Revenue+ logAverage_Block_Size, data=train)
summary(mod4)

vif(mod4)
which(cooks.distance(mod4)>qf(0.5, 4, 373-4))
which(abs(dffits(mod4)) > 2*sqrt(4/373))
par(mfrow=c(1,1))
plot(rstandard(mod4)~train$logMiner.s_Revenue)
plot(rstandard(mod4)~train$logGold_Price)
plot(rstandard(mod4)~train$Average_Block_Size)
qqnorm(rstandard(mod4))
qqline(rstandard(mod4))

```


```{r}
mod5 <- lm(logBTC_Price ~ Transaction_Vol+ logMiner.s_Revenue + Exchange_Vol + logGold_Price + logAverage_Block_Size, data=train)
summary(mod5)

vif(mod5)
which(cooks.distance(mod5)>qf(0.5, 6, 373-6))
which(abs(dffits(mod5)) > 2*sqrt(6/373))
par(mfrow=c(1,1))
plot(rstandard(mod5)~train$Transaction_Vol)
plot(rstandard(mod5)~train$logMiner.s_Revenue)
plot(rstandard(mod5)~train$Exchange_Vol)
plot(rstandard(mod5)~train$logGold_Price)
plot(rstandard(mod5)~train$Average_Block_Size)
qqnorm(rstandard(mod5))
qqline(rstandard(mod5))

vif(mod5)
```

```{r}
full2test <- lm(logBTC_Price ~ ., data=test[,-c(1,2,3,4,5,8,9,10)])
mod3test <- lm(logBTC_Price ~ logETH._Price + logRipple_Price + logMiner.s_Revenue + logGold_Price + logMSCI_World + logAverage_Block_Size, data=test)
mod4test <- lm(logBTC_Price ~ logGold_Price + logMiner.s_Revenue + logAverage_Block_Size, data=test)
mod5test<-lm(logBTC_Price ~ Transaction_Vol+ logMiner.s_Revenue + Exchange_Vol + logGold_Price + logAverage_Block_Size, data=test)
```

```{r}
# compare full2 's
summary(full2)
summary(full2test)
vif(full2test)
which(cooks.distance(mod3test)>qf(0.5, 9, 373-9))
which(abs(dffits(mod3test)) > 2*sqrt(9/373))
par(mfrow=c(1,1))
plot(rstandard(mod3)~test$logETH._Price)
plot(rstandard(mod3)~test$logRipple_Price)
plot(rstandard(mod3)~test$logMiner.s_Revenue)
plot(rstandard(mod3)~test$logGold_Price)
plot(rstandard(mod3)~test$logMSCI_World)
plot(rstandard(mod3)~test$logAverage_Block_Size)
plot(rstandard(mod3)~test$Exchange_Vol)
plot(rstandard(mod3)~test$Transaction_Vol)

qqnorm(rstandard(full2test))
qqline(rstandard(full2test))
```


```{r}
# compare model 3's
summary(mod3)
summary(mod3test)
vif(mod3test)
which(cooks.distance(mod3test)>qf(0.5, 7, 373-7))
which(abs(dffits(mod3test)) > 2*sqrt(7/373))
par(mfrow=c(1,1))
plot(rstandard(mod3)~test$logETH._Price)
plot(rstandard(mod3)~test$logRipple_Price)
plot(rstandard(mod3)~test$logMiner.s_Revenue)
plot(rstandard(mod3)~test$logGold_Price)
plot(rstandard(mod3)~test$logMSCI_World)
plot(rstandard(mod3)~test$Average_Block_Size)
qqnorm(rstandard(mod3test))
qqline(rstandard(mod3test))
```


```{r}
# compare model 4's
summary(mod4)
summary(mod4test)
vif(mod4test)
which(cooks.distance(mod4test)>qf(0.5, 4, 373-4))
which(abs(dffits(mod4test)) > 2*sqrt(4/373))
par(mfrow=c(1,1))
plot(rstandard(mod4)~test$logMiner.s_Revenue)
plot(rstandard(mod4)~test$logGold_Price)
plot(rstandard(mod4)~test$Average_Block_Size)
qqnorm(rstandard(mod4test))
qqline(rstandard(mod4test))
```
```{r}
# compare model 5's
summary(mod5)
summary(mod5test)
vif(mod5test)
which(cooks.distance(mod5test)>qf(0.5, 7, 373-7))
which(abs(dffits(mod5test)) > 2*sqrt(7/373))
par(mfrow=c(1,1))
plot(rstandard(mod5)~test$Transaction_Vol)
plot(rstandard(mod5)~test$logMiner.s_Revenue)
plot(rstandard(mod5)~test$Exchange_Vol)
plot(rstandard(mod5)~test$logGold_Price)
plot(rstandard(mod5)~test$Average_Block_Size)
qqnorm(rstandard(mod5test))
qqline(rstandard(mod5test))

```




